On Logic Embeddings and G\"odel's God

### 2015-03-02 Papers/2014/WADT/WADT-4pages.tex
slightly modified WADT paper

We briefly summarize some of these observations: Nitpick confirms that
Scott's axioms are consistent, while LEO-II and Satallax demonstrate that
Gödel's original, slightly different axioms are inconsistent. As far
as we are aware, this is a new result. As experiments with LEO-II
revealed, the problem lies in a subtle difference in the definitions
of the predicate \textit{essence} (characterizing the essential
properties of an entity) between Gödel and Scott. In several philosophy papers\marginpar{citation needed? weaken the "several" claim?}, however, this
difference has been commented as a simple cosmetic change.  Moreover,
instead of using modal logic S5, 
LEO-II and Satallax can prove the final theorem (that is,
$\square \exists x . G(x)$, necessarily there exists God) already for
modal logic KB.  This is highly relevant since several philosophers
have criticized G{\"o}del's argument for the use of logic S5\marginpar{citation needed? weaken the "several" claim?}.  Axiom
B (symmetry), however, cannot be dropped, which in turn is confirmed by Nitpick.
LEO-II and Satallax can also show that G{\"o}del's and Scott's axioms
imply what is called the modal collapse: $\phi\supset\Box\phi$. This
expresses that contingent truth implies necessary truth (which
can even be interpreted as an argument against free will;
cf.~\cite{sobel2004logic}) and is probably the most
fundamental criticism put forward against G{\"o}del's and Scott's
versions of the argument. Other theorems that can be shown by LEO-II
and Satallax include flawlessness of God and monotheism.

>>>>

We briefly summarize some of these observations: Nitpick confirms that
Scott's axioms are consistent, while LEO-II and Satallax demonstrate
that Gödel's original, slightly different axioms are inconsistent. As
far as we are aware, this is a new result. As experiments with LEO-II
revealed, the problem lies in a subtle difference in the definitions
of the predicate \textit{essence} (characterizing the essential
properties of an entity) between Gödel and Scott. In recent papers on
the ontological argument (see e.g. below), some authors speak of an oversight/flaw by
Gödel, some silently replace Gödel's definition without commenting and
some simply stay with it.
Moreover, instead of using modal logic S5, LEO-II and Satallax can
prove the final theorem (that is, $\square \exists x . G(x)$,
necessarily there exists God) already for modal logic KB.  This is
highly relevant since some philosophers have criticized G{\"o}del's
argument for the use of logic S5.  Axiom B (symmetry), however, cannot be dropped,
which in turn is confirmed by Nitpick.  LEO-II and Satallax can also
show that G{\"o}del's and Scott's axioms imply what is called the
modal collapse: $\phi\supset\Box\phi$. This expresses that contingent
truth implies necessary truth (which can even be interpreted as an
argument against free will; cf.~\cite{sobel2004logic}) and is probably
the most fundamental criticism put forward against G{\"o}del's and
Scott's versions of the argument. Other theorems that can be shown by
LEO-II and Satallax include flawlessness of God and 
monotheism.
